Tracking and predicting turn structure during language acquisition

### 2014-09-19 paper/CasillasFrank-Anticip.tex
Some changes to text, mostly in discussion. Figure axis changes. Updates to references.

\author[MPI]{Marisa Casillas}
% \address[StanfordLX]{Department of Linguistics, Stanford University}
\address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}

@@

\begin{abstract}
Blah blah blah abstract needs to be written here.

\end{abstract}

@@

Spontaneous conversation is a universal context for using and learning language. Like other types of human interaction, it is organized at its core by the roles and goals of its participants. What sets conversation apart is its structure, however: sequences of interconnected, communicative actions that take place across alternating turns at talk. Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities. Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{stivers2009, dingemanse2013, de-vosInPrep}. How does this ability develop?

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep{snow1977, jaffe2001, hilbrinkInPrep}. During their first year, infants know little about the lexical and syntactic systems of their language \citep[though they \textit{do} know a little about prosody;][]{christophe2001, bertoncini2013}. Of course, infant turn-taking is different from adult turn-taking in several ways. Infant turn-taking is heavily scaffolded by caregivers, has timing distinct from adult turn-taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001, snow1977}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn-taking. Early non-verbal interactions establish the protocol by which children come to use language with others.

% If we want to grasp children's linguistic input and practice, we need to understand their conversational skill development. And if we want to understand how children develop conversational skills, we need to take a close look at how turn-taking develops in tandem with language. 
In this study, we investigate when children begin to make predictions about upcoming turn structure, and how they integrate language into their predictions as they grow older. In what follows, we first review background about turn-taking and the state of current knowledge about adult turn prediction; we then discuss recent work on the development of turn-taking before turning to the details of our own study.

% For example, they show sensitivity to language-specific aspects of prosody in the first few days of life, and can use prosody for word segmentation and emotion recognition by 7--8 months \citep{jusczyk1999, johnson2001, grossmann2005}. In contrast, children don't know much about syntactic structure until their third year \citep{clark2009}.

@@

Turn-taking itself is not unique to conversation. Many other human activities are organized around sequential turns. Traffic in intersections and communication in computer networks both have characteristic turn-taking. Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}. Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}. In all these cases, turn-taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.

Conversation distinguishes itself from non-conversational turn-taking behaviors, however, by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what types of talk will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).

Despite this complexity, conversational turn-taking is precise in its timing. Across a diverse sample of 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}. Current models of speech production suggest that it takes approximately 600 msec to produce a word, and even longer to produce a simple utterance \citep{levelt1989}. So, in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended. To formulate their response early on, speakers must track and anticipate what types of response might become relevant next. They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment. Prediction thus plays a key role in timely turn-taking.

@@

Adults have a lot of information that could potentially help with their predictions about upcoming turn structure. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, bogelsUndRev}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}. Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information is plays more of a secondary role \citep{de-ruiter2006}. For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, bogelsUndRev}).

Recent experimental work on adult turn-taking has tried to tease apart which linguistic cues responders rely on in making online predictions about upcoming turn structure. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and press a button whenever they anticipated that the current speaker was about to finish her turn. The snippets that participants listened to were controlled for the amount of linguistic information present; some snippets were normal, unmodified versions of the recording, but some had words and rhythm without pitch, pitch and rhythm but no words, neither words nor pitch, or were replaced entirely with a noise profile matched to the original recording. In the normal, unmodified snippets, participants' anticipatory button presses were highly accurate, falling an average of 180 msec before the phonetic end of the turns. Participants also made accurate predictions for the ``words and rhythm without pitch'' snippets; there was no significant difference compared to their performance in the normal  condition. But, in the snippets with no lexical information, participants' accuracy significantly declined. De Ruiter and colleagues \citeyearpar{de-ruiter2006} concluded that lexicosyntactic information is the primary linguistic signal listeners use for making online predictions about turn-end boundaries. They also found a significant, but secondary role for pitch: participants' accuracy in the pitch and rhythm (but no words) condition was still better than the neither-words-nor-pitch and the noise profile conditions. They concluded that pitch's importance for online prediction of turn-ends is restricted, and secondary to lexicosyntactic information.

Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content. 
% Accurate prediction of linguistic and pragmatic content is essential to the mechanisms underlying adult turn taking. 
Adults anticipate turn ends better when they can more accurately predict exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}). They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible. \citet*{bogelsmagyariInPrep} used a trivia set-up to get participants to answer questions in which the answer became known either early or late in the question (e.g., ``Which character, also called 007, appears in the famous movies?'' vs. ``Which character from the famous movies, is also called 007?''). Their participants showed a consistent neural response 500 msec after the answer became unambiguous, suggesting immediate response planning.

In another study investigating the relationship between lexicosyntactic and prosodic cues, \citet*{bogelsUndRev} used a combination of button-press and verbal responses.
 % to more explicitly test the role of prosodic cues in turn-boundary identification. 
They found that participants never verbally responded or pressed the ``anticipation'' button when hearing a syntactically-complete unit without a boundary tone. But utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525). Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

In sum, adults accurately and spontaneously make predictions about upcoming turn structure. They use a wide variety of cues in making their predictions and then act immediately upon them. Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions. The acquisition of turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues. But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues thus doesn't fit well with children's pre-verbal turn-taking.

@@

The majority of work on children's early turn-taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{snow1977, jaffe2001}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}. Infants at this age also notice changes to social contingency outside of turn structure. In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze. When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. Infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency. But then children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.

@@

There are mixed results regarding when children begin to anticipate turn structure in conversation. One study found that 12-month-olds make more predictive gaze shifts while watching human verbal conversation compared to object non-speech ``conversation'' \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}. Neither of these studies had baselines to which the turn-relevant looking behavior could be compared, however. Such a baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if such shifting is not related to turn structure. These developmental differences could produce artifactual changes in measures of turn-contingent shifting.

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds, finding that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only three-year-olds were affected by a lack of intonation. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 3;0), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 3;0).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it nevertheless has several limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline gaze frequency were computed only during conversational turns (not pauses), while their estimates of anticipatory shifting were computed during pauses. Assuming that gaze switches happen most often during the gap between turns \citep{hirvenkari2013}, their choice of a ``random'' (baseline) gaze window would have maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
% \footnote{If so, the null effect for children under 3;0 could be given a stronger interpretation and the significant effect for three-year-olds and adults could be given a weaker interpretation.} 
A stronger baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus. We follow this approach in our work.

Second, the conversation stimuli they used were somewhat unusual. The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps are often around 200 msec \citep{stivers2009}. The speakers in the video were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus. In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers. We again attempt to address these issues in our current study.

% A single-channel recording was most problematic for the no-intonation condition, where the two voices were flattened to have the same pitch. 
Despite these minor methodological issues, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and prosodic development \textit{do} play a role in children's looking patterns. Our current work thus takes this paradigm as our starting point.

@@

We report here on the role of linguistic processing in children's predictions about upcoming turn structure. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available. Experiment 1 uses a paradigm in which participants view video clips of naturalistic conversations across a number of different languages, so as to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were relatively constant. This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we created artificial (puppet) visual stimuli to better separate these cues from one another. In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.

In both experiments, children consistently looked faster to responders after hearing questions, compared to non-questions. Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages. Our findings overall support an account in which predictive processes for turn-taking in conversation are present early but their integration with linguistic information takes substantial practice.

@@

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See the videos here: NOTE: add link.}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
\end{center}
\caption{Baseline-corrected proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:conditionsFL}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:questionsFL}
\end{figure}

@@

Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56). As predicted, children made more anticipatory looks while they were watching videos that they could understand. In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure. Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.

Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145). Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063). We had expected that, as children got older, they would also show more anticipatory switches. But this is contradicted by decreased looks with age for English videos, which even extended to the adult data. One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely. Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations. In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).

>>>>

\author[MPI]{Marisa Casillas\corref{cor1}}
% \address[StanfordLX]{Department of Linguistics, Stanford University}
\address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}
\cortext[cor1]{Corresponding author}

@@

\begin{abstract}
We investigate language acquisition through the lens of a fundamental conversational skill: turn taking. Children begin developing turn-taking skills in infancy, but take several years to assimilate language into their turn-taking systems. In two eye-tracking experiments with children ages 1--6, we measured participants' spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available. We found that children already predicted upcoming turns at age one, and integrated linguistic cues differently at different ages. Children under 3;0 showed a general advantage for prosody over lexicosyntax, contrary to the typical finding for adults. Children 2;0 and older showed more predictive switches for questions than non-questions, but only when lexical information was available. But, we found no evidence that lexicosyntax is sufficient for turn prediction---participants' performance was best overall with access to lexicosyntax and prosody together. Our results point to the importance of interaction in children’s real-time linguistic processing.
\end{abstract}

@@

Spontaneous conversation is a universal context for using and learning language. Like other types of human interaction, it is organized at its core by the roles and goals of its participants. But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk. Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities. Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{stivers2009, dingemanse2013, de-vosInPrep}. How does this ability develop?

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep{snow1977, jaffe2001, hilbrinkInPrep}. Of course, infant turn-taking is different from adult turn-taking in several ways. Infant turn-taking is heavily scaffolded by caregivers, has timing distinct from adult turn-taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001, snow1977}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn-taking. Early non-verbal interactions establish the protocol by which children come to use language with others. And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.

In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation, and how they integrate language into their predictions as they grow older. In what follows, we first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction. We then discuss recent work on the development of turn-taking skills before turning to the details of our own study.

@@

Turn-taking itself is not unique to conversation. Many other human activities are organized around sequential turns. Traffic intersections and computer network communication both use turn-taking systems. Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}. Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}. In all these cases, turn-taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what types of talk will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).

Despite this complexity, conversational turn-taking is precise in its timing. Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}. Current models of speech production suggest that it takes approximately 600 msec to produce a word, and even longer to produce a simple utterance \citep{levelt1989}. So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended. To formulate their response early on, speakers must track and anticipate what types of response might become relevant next. They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment. Prediction thus plays a key role in timely turn-taking.

@@

Adults have a lot of information that could potentially help with their predictions about upcoming turn structure. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}. Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information plays more of a secondary role \citep{de-ruiter2006}. For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, torreiraUndRev}).

Recent experimental work on adult turn-taking has tried to tease apart which linguistic cues responders rely on in making online predictions about upcoming turn structure. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish her turn. The snippets that participants listened to were controlled for the amount of linguistic information present; some snippets were normal, unmodified versions of the recording, but some had words and rhythm (but no pitch), pitch and rhythm (but no words), neither words nor pitch, or were replaced entirely with a noise profile matched to the original recording. In the normal, unmodified snippets, participants' anticipatory button presses were highly accurate, falling an average of 180 msec before the phonetic end of the turns. Participants also made accurate predictions for the snippets with words and rhythm (but no pitch); there was no significant difference compared to their performance in the normal condition. But, in the snippets with no lexical information, participants' accuracy significantly declined. De Ruiter and colleagues \citeyearpar{de-ruiter2006} concluded that lexicosyntactic information is the primary linguistic signal listeners use for making online predictions about turn-end boundaries. They also found a significant, but secondary role for pitch: participants' accuracy in the pitch and rhythm (but no words) condition was still better than the neither-words-nor-pitch and the noise profile conditions. They concluded that pitch's importance for online prediction of turn-ends is restricted, falling secondary to lexicosyntactic information.

Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content. Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}). They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible. \citet*{bogelsmagyariInPrep} used a trivia quiz set-up to get participants to answer questions in which the answer became known either early or late in the question (e.g., ``Which character, also called 007, appears in the famous movies?'' vs. ``Which character, from the famous movies, is also called 007?''). Their participants showed a consistent neural response 500 msec after the answer became unambiguous, suggesting immediate response planning.

In another study, \citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-boundary identification. They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without a boundary tone. However, utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525). Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

In sum, adults accurately and spontaneously make predictions about upcoming turn structure. They use a wide variety of cues in making their predictions and then act immediately upon them. Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions. The acquisition of turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues. But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn-taking.

@@

The majority of work on children's early turn taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{snow1977, jaffe2001}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}. Infants at this age also notice changes to social contingency outside of turn structure. In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze. When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. Infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency. But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.

@@

There are mixed results regarding when children begin to anticipate turn structure in conversation. One study found that 12-month-olds make more predictive gaze shifts while watching human verbal conversation compared to object non-speech ``conversation'' \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}. However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared. A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure. Such developmental differences could produce artifactual changes in measures of turn-contingent shifting.

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds, finding that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only 36-month-olds were affected by a lack of intonation. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 3;0), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 3;0).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it nevertheless has several limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline ``random'' gaze frequency were computed from switches that occurred while the speaker's turn was ongoing (i.e, during turns, and not during inter-turn gaps). Their estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps (i.e., around inter-turn gaps, and primarily not during turns). Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, switches made during ongoing speech (their ``random'' baseline frequency) would be at a minimum. Therefore their ``random'' (baseline) frequency actually maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies. A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus. We follow this latter approach in our work.

Second, the conversation stimuli they used were somewhat unusual. The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}. The speakers in the videos were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus. In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers. Again, we attempt to address these issues in our current study.

Despite these minor methodological issues, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and prosodic development \textit{do} play a role in children's looking patterns. Our current work thus takes this paradigm as our starting point.

@@

We report here on the role of linguistic processing in children's predictions about upcoming turn structure. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available. Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages, to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact. This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we created artificial (puppet) visual stimuli to better separate these linguistic cues from one another. In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.

In both experiments children consistently looked faster to responders after hearing questions, compared to non-questions. Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages. Our findings overall support an account in which predictive processes for turn-taking in conversation are present early, but their integration with linguistic information takes substantial practice.

@@

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See the videos here: NOTE: Link to be added.}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
\end{center}
\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:conditionsFL}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
\end{center}
\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:questionsFL}
\end{figure}

@@

Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} and \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand. In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure. Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.

Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145, \textit{p}=.252). Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063, \textit{p}$<$.001). We had expected that, as children got older, they would also show more anticipatory switches. But this is contradicted by decreased looks with age for English videos, which even extended to the adult data. One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely. Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations. In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).
